---
title: "p8105_hw3_qz2266"
author: "Qing Zhou"
date: "2022-10-15"
output: github_document
---


```{r, message = FALSE}
library(tidyverse)
library(lubridate)
library(readxl)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
```


### Problem 2

#### Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset. 

```{r}
accel = 
  read_csv("data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "act_minute",
    values_to = "act_count") %>%
 
  mutate(
    day_type = case_when(
      day == "Monday" ~ "weekday",
      day == "Tuesday" ~ "weekday",
      day == "Wednesday" ~ "weekday",
      day == "Thursday" ~ "weekday",
      day == "Friday" ~ "weekday",
      day == "Saturday" ~ "weekend",
      day == "Sunday" ~ "weekend" ,
      TRUE     ~ ""
  )) %>%
  
  mutate(
    week = as.integer(week),
    day_id = as.integer(day_id),
    day = factor(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")),
    day_type = factor(day_type, levels = c("weekday", "weekend"))) %>% 
  relocate(week,day_id, day, day_type)
```

The data set has `r nrow(accel)` observations, and contains `r ncol(accel)` variables, including `r colnames(accel)`. Data was collected across `r accel %>% summarise(count = n_distinct(day_id))` days of `r accel %>% summarise(count = n_distinct(week))` weeks.

There are four date variables :  

* day_id is a `r class(pull(accel,day_id))` variable that indicates the day when data was collected;
* day is a `r class(pull(accel,day))` variable  that indicates which day it is when data was collected;
* week is a `r class(pull(accel, week))` variable that indicates the week when data was collected.
* day_type is a `r class(pull(accel, day_type))` variable that indicates the information was collected on weekend or weekday.

There are two activity variables :

* act_minute is a `r class(pull(accel, act_minute))` variable that indicates the minute when the activity count was collected, corresponding to each minute of a 24-hour day starting at midnight.
* On each day, there are `r accel %>% filter(day_id == 1) %>% summarise(count = n_distinct(act_minute))` activity counts for each minute of a 24-hour day starting at midnight, and the corresponding value is stored in act_count, which is a `r class(pull(accel, act_count))` variable.


#### Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals.

```{r   total activity}
library(dplyr, warn.conflicts = FALSE)

options(dplyr.summarise.inform = FALSE)

total_act =
  accel %>%
  group_by(week, day) %>%
  dplyr::summarise(total_activity = sum(act_count)) %>% 
  pivot_wider(
    names_from = "day", 
    values_from = "total_activity")

```

We couldn't see any apparent trends over days. However, we could roughly find that the total activity across 5 weeks is high on Wednesday, Thursday and Friday, and low in Saturday. The total activity seemed to peak on Friday but quickly drop to the bottom on Saturday.  

More specifically, the total activity sum of Sunday is `r sum(total_act[, 'Sunday'])`, of Saturday is `r sum(total_act[, 'Saturday'])`, of Monday is `r sum(total_act[, 'Monday'])`, of Tuesday is `r sum(total_act[, 'Tuesday'])`, of Wednesday is `r sum(total_act[, 'Wednesday'])`, of Thursday is `r sum(total_act[, 'Thursday'])`, and of Friday is `r sum(total_act[, 'Friday'])`.

The user is relatively more active at the beginning of the study than later on. On Saturdays of week 4 and 5, the data were much lower than the rest. There might be some problem with the measurement, such as incorrect use of the device. 


#### Make a single-panel plot to show the 24-hour activity time courses for each day.

```{r   accel_plot}
accel_plot = 
  accel %>%
  separate(act_minute, into = c("activity", "act_min")) %>%
  select(-activity) %>%
  mutate(
    act_min = as.numeric(act_min))

accel_plot %>%
  group_by(day_id) %>%
  ggplot(aes(x = act_min, y = act_count, color = day)) +
  geom_point() +
  geom_line() +
  
  labs(
    title = "24-hour Activity Count by Day",
    x = "Minute",
    y = "Activity Count"
  )
```

#### To make it easier to intepret the data, we use the average activity at each minute of the same day of 5 weeks as the y-axis and time points in a 24-hour pattern as the x-axis:

```{r   accel_mean_plot}
accel_mean_plot = 
  accel %>%
  separate(act_minute, into = c("activity", "act_min")) %>%
  select(-activity) %>%
  mutate(
    act_min = as.numeric(act_min)) %>%
  group_by(day, act_min) %>% 
  dplyr::summarise(avg_count = mean(act_count))

accel_mean_plot %>%
  ggplot(aes(x = act_min, y = avg_count, color = day)) + 
  geom_point() +
  geom_line(alpha = 0.5) +
  labs(
    title = "24-hour Average Activity Count by Day",
    x = "Time Course",
    y = "Average Activity Count"
  ) + 
  scale_x_continuous(
    breaks = c(0, 360, 720, 1080, 1440), 
    labels = c("12AM", "6AM", "12PM", "6PM", "11:59PM"),
    limits = c(0, 1440)
    )
```

From the plot we could see the average activity count at most time points are below 2500.There are three peaks located around 7-8AM, 12PM and 8-9pm. Especially on Friday, there is an obvious peak around 8-9pm. Moreover, at nights from 12AM to 5AM, the activity counts are pretty low. It might be because this time slot is the sleeping hour.



### Problem 3

#### Write a description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue.

```{r  load data}
library(p8105.datasets)
data("ny_noaa")
```

```{r  summary}
ny_noaa %>%
  group_by(id) %>%
  summarize(n_obs = n()) 

ny_noaa %>%
  summary() 
```

```{r  missing data, results = 'hide'}
# number of rows with missing data
ny_noaa %>% drop_na() %>% nrow()

# number of columns with missing data
sum(is.na(ny_noaa$id))
sum(is.na(ny_noaa$date))
sum(is.na(ny_noaa$prcp))
sum(is.na(ny_noaa$snow))
sum(is.na(ny_noaa$snwd))
sum(is.na(ny_noaa$tmax))
sum(is.na(ny_noaa$tmin))
```

The ny_noaa data set has `r nrow(ny_noaa)` observations, and  `r ncol(ny_noaa)` variables, including `r colnames(ny_noaa)`:

id: Weather station ID
date: Date of observation
prcp: Precipitation (tenths of mm)
snow: Snowfall (mm)
snwd: Snow depth (mm)
tmax: Maximum temperature (tenths of degrees C)
tmin: Minimum temperature (tenths of degrees C)

Data was collected from 747 weather stations, ranging from 1981-01-01 to 2010-12-31.

There are `r ny_noaa %>% drop_na() %>% nrow()` rows with missing data. Among them, `r sum(is.na(ny_noaa$prcp))/nrow(ny_noaa)*100` % missing data for precipitation variable,  
`r sum(is.na(ny_noaa$snow))/nrow(ny_noaa)*100` % for snowfall,  `r sum(is.na(ny_noaa$snwd))/nrow(ny_noaa)*100` % for snow depth, `r sum(is.na(ny_noaa$tmax))/nrow(ny_noaa)*100` % for maximum temperature, and  `r sum(is.na(ny_noaa$tmin))/nrow(ny_noaa)*100` % for minimum temperature.
Therefore, the missing data is an issue, especially for tmax and tmin.


#### Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values?

```{r}
ny_noaa =
ny_noaa %>% 
  mutate(
  year = lubridate::year(date),
  month = lubridate::month(date),
  day = lubridate::day(date),
  prcp = prcp/10,
  tmin = as.numeric(tmin)/10,
  tmax = as.numeric(tmax)/10
  ) %>% 
  relocate(id, year, month, day) %>% 
  select(-date)
```

```{r}
ncol(ny_noaa)
colnames(ny_noaa)
```

```{r snow_value}
ny_noaa %>%
  count(snow) %>%
  arrange(desc(n)) %>%
  top_n(3) 
```

The most commonly observed values for snowfall are 0, NA and 25.

#### Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers? 

```{r ny_noaa_plot1}
jan_july = 
  ny_noaa %>%
  mutate(
    month = recode_factor(month,
          "01" = "January",
          "02" = "February",
          "03" = "March",
          "04" = "April",
          "05" = "May",
          "06" = "June",
          "07" = "July",
          "08" = "August",
          "09" = "September",
          "10" = "October",
          "11" = "November",
          "12" = "December"),
    year = as.numeric(year)) %>%
  filter(month %in% c("January", "July"))  %>%
  select(id, month, year, tmax) %>%
  group_by(id, year, month) %>% 
  dplyr::summarise(mean_tmax = mean(tmax, na.rm = TRUE))

summary(jan_july)
```

```{r}
jan_july %>% ggplot(aes(x = year, y = mean_tmax, color = month)) +
  geom_point(alpha = 0.5) + 
  geom_smooth() + 
  labs(title = "Average Max Temperature in January and July ",
    x = "Year",
    y = "Average Max Temperature (C)" ) +
  facet_grid(~month) 
```

**Comments**: Between 1981 and 2010, the average maximum temperature in January is lower than the one in July (about 0 vs. 25 degree celsius), which is consistent with the change of seasons in NY. Moreover, generally speaking, the average maximum temperature in January of each year seems vary stronger than the one in July, across all the weather stations, but their overall trends are quite similar to each other. As for the outliers, most of them were located in both positive and negative direction in January, while those in July appeared mostly in the negative direction. 


#### Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year

```{r tmax_tmin}
tmax_tmin_p = 
  ny_noaa %>% 
  filter(tmax != "NA" & tmin != "NA") %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex(bin = 20) +
  geom_smooth(se = F) +
  scale_fill_viridis_c() +
  labs(
    title = "NY Max and Min Temperatures (1981-2010)",
    x = "Max Temp (째C)",
    y = "Min Temp (째C)"
  )
  theme(legend.position = "right")
  
tmax_tmin_p
```

```{r snowfall}
snowfall_p = 
  ny_noaa %>% 
  drop_na(snow) %>%
  filter(snow > 0, snow < 100) %>% 
  mutate(
    year =  as.factor(year)
    ) %>%
  ggplot(aes(x = year, y = snow)) +
  geom_boxplot(fill = '#A4A4A4', color = "blue") +
  
  labs(
      x = "year",
      y = "snowfall (mm)",
      title = "NY Snowfall Distribution (1981-2010)"
    ) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, size = 5))

snowfall_p
```

```{r patchwork}
library(patchwork)
tmax_tmin_p / snowfall_p
```

From the above NY Max and Min Temperatures plot we can see from 1981-2010, the temperature differences in NY were not big. Basically, when the max temperature goes up, the min temperature goes up as well and verse visa. The line represents the relationship between the max temperature and the min temperature. It is roughly linear and most days fluctuates around this line. 

There were very few days the temperature went to extreme, since the most days the max temperatures were below 40 and the min temperatures were above -30 degree celsius.The most frequently appeared maximum temperature is ~25째C and the most frequently appeared minimum temperature is ~15째C.

From the below NY Snowfall Distribution plot we can see the distribution of snowfall values that are greater than 0 and less than 100 separately by year, from 1981-2010. The snowfall remains quite even across the entire period, with a right-skewed distribution or long tail towards upper snowfall amounts. The median snowfall is around 25 mm and stays consistent year to year, give or take a few mm. The upper and lower quartiles are also quite consistent, with the 25th percentile around 15 mm and 75th percentile around 50 mm. There are more outliers showing up in the year of 1998, 2006 and 2010.

### Problem 1

#### Read in the data
```{r}
library(ggridges)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```


